Cradles of the first stars: self-shielding, halo masses, and multiplicity

### 2019-08-05 paper.tex
Fixed some formatting. Added notes to referee report and fixed a few different things within the paper (all in bold).

We simulate a 1 Mpc$^{3}$ comoving box with a 256$^{3}$ base grid 
resolution and a dark matter paticle mass of 2001 M$_{\odot}$. It has a maximum refinement level of 12 which provides a maximal comoving resolution of $\sim$1 pc. The simulation is 
initialized with \music{} \citep{Hahn11_MUSIC} at $z = 130$ and uses the cosmological parameters from the Planck collaboration best fit 
\citet{Planck13_Cosmo}: $\Omega_{\rm M}$ = 0.3175, $\Omega_{\Lambda}$ = 
0.6825, $\Omega_{\rm DM}$ = 0.2685, and $h = 0.6711$, with the variables 
having their usual definitions. The simulation is run until $z = 9.32$, when it becomes too computationally expensive to continue. At this point, about 20\% of the volume is over 10\% ionized. We output 918 datasets,  roughly 0.5 Myr apart. In this paper, we focus only on outputs where Pop III stars have just formed, from $z$ = 27.23 to 9.39.

@@

If within 1 pc, multiple cells meet this criteria, then a single Pop III star forms at the center of mass of these cells. The mass of the Pop III star is randomly sampled from an initial mass function (IMF) of the form:
\begin{equation} \label{IMF}
	f(\log M)dM = M^{-1.3} \exp \left[-\left( \frac{M_{\rm char}}{M}\right)^{1.6} \right]dM
\end{equation}
where M$_{\rm char}$ = 20 M$_{\odot}$ \citep{Hirano17}. This characteristic mass is different from the W12 choice of 100 M$_{\odot}$. This results in an exponential cutoff below M$_{\rm char}$ and a power-law IMF above. Once the cell meets the above criteria and the stellar mass is chosen, an equal amount of gas is removed from the grid in a sphere which contains twice the stellar mass that is centered on the new star particle. The new Pop III star then gains the mass-weighted velocity of the gas contained within that sphere. In this simulation, we do not track any small scale fragmentation which may form low mass stars \citep{Greif11_P3Cluster}. We are only tracking massive Pop III star formation.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/shield_mass.pdf}
    \caption{Shielding factor as a function of halo mass. Solid lines indicate a core radius of $f_{c} = 10^{-2}$ and the dashed lines indicate a core radius of $f_{c} = 10^{-4}$. For these parameters, and assuming the Doppler broadening parameter equals the circular velocity, the shielding factor is below 1 percent. The green line shows the shielding factor as a function of \hh{} column density for a circular velocity of 10 km/s. }
    \label{fig:shield_mass}
\end{figure}
%====================================================================
\section{Results}
%====================================================================
In this section, we will present the distribution of host halo masses of Pop III stars, the relation between the mass distribution and the LW background radiation, and the distribution of various Pop III properties.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/hmf.pdf}
    \caption{Halo mass function of the last output of the simulation at $z$ = 9.3. The thick, black line is the analytic Sheth-Tormen mass function. The distribution of halo masses matches well with analytical expectations above 10$^{5}$ M$_{\odot}$, where halos are captured by 50 particles.}
    \label{fig:hmf}
\end{figure}

@@

%====================================================================
\subsection{Population III host halo masses}
%====================================================================
 To determine the host halos of new Pop III stars, we use the halo finding code, \rockstar{} \citep{rockstar}. The mass determined by \rockstar{} is the dark matter mass, so to determine the total mass of the halos, we multiply the masses by $\Omega_{\rm M}/\Omega_{\rm DM}$. The total mass is subsequently used in our analysis. We also take the center and virial radius from the \rockstar{} halo catalog. A small percentage of Pop III stars do not form in a halo identified by \rockstar{}. Assuming that Pop III stars must form in collapsed halos, we calculate the virial mass and radius of a halo centered on the Pop III star directly from the simulation data by determining the sphere that contains an average overdensity of 200 $\rho_{\rm c}$. We calculate the LW intensity impinging on each halo by summing up the contribution coming from each radiating star particle outside the virial radius of the host halo. This is then added to the constant LW background implemented in the simulation (Eq. \ref{LWbg}). In this work, we only analyze halos below the atomic cooling limit.

>>>>

We simulate a 1 Mpc$^{3}$ comoving box with a 256$^{3}$ base grid 
resolution and a dark matter paticle mass of 2001 M$_{\odot}$. It has a maximum refinement level of 12 which provides a maximal comoving resolution of $\sim$1 pc. The simulation is 
initialized with \music{} \citep{Hahn11_MUSIC} at $z = 130$ and uses the cosmological parameters from the Planck collaboration best fit 
\citet{Planck13_Cosmo}: $\Omega_{\rm M}$ = 0.3175, $\Omega_{\Lambda}$ = 
0.6825, $\Omega_{\rm DM}$ = 0.2685, and $h = 0.6711$, with the variables 
having their usual definitions. The simulation is run until $z = 9.32$, when it becomes too computationally expensive to continue. At this point, about 20\% of the volume is over 10\% ionized. We output 918 datasets,  roughly 0.5 Myr apart. In this paper, we focus only on outputs where Pop III stars have just formed, from $z$ = 27.23 to 9.39. \textbf{Throughout the rest of the paper, all numbers are in physical units.}

@@

If within 1 pc, multiple cells meet this criteria, then a single Pop III star forms at the center of mass of these cells. The mass of the Pop III star is randomly sampled from an initial mass function (IMF) of the form:
\begin{equation} \label{IMF}
	f(\log M)dM = M^{-1.3} \exp \left[-\left( \frac{M_{\rm char}}{M}\right)^{1.6} \right]dM
\end{equation}
where M$_{\rm char}$ = 20 M$_{\odot}$ \citep{Hirano17}. This characteristic mass is different from the W12 choice of 100 M$_{\odot}$. This results in an exponential cutoff below M$_{\rm char}$ and a power-law IMF above. \textbf{The Pop III IMF mass ranges from $1 \leq M / M_{\odot} \leq 300$. }Once the cell meets the above criteria and the stellar mass is chosen, an equal amount of gas is removed from the grid in a sphere which contains twice the stellar mass that is centered on the new star particle. The new Pop III star then gains the mass-weighted velocity of the gas contained within that sphere. In this simulation, we do not track any small scale fragmentation which may form low mass stars \citep{Greif11_P3Cluster}. We are only tracking massive Pop III star formation.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/shield_mass.pdf}
    \caption{Shielding factor as a function of halo mass. Solid lines indicate a core radius of $f_{c} = 10^{-2}$ and the dashed lines indicate a core radius of $f_{c} = 10^{-4}$. For these parameters, and assuming the Doppler broadening parameter equals the circular velocity, the shielding factor is below 1 percent. The green line shows the shielding factor as a function of \hh{} column density for a circular velocity of 10 km/s. \textbf{Note that a shielding factor of 1 means that there is no shielding and a shielding factor of 0 means complete shielding.}}
    \label{fig:shield_mass}
\end{figure}
%====================================================================
\section{Results}
%====================================================================
In this section, we will present the distribution of host halo masses of Pop III stars, the relation between the mass distribution and the LW background radiation, and the distribution of various Pop III properties.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/hmf.pdf}
    \caption{Halo mass function of the last output of the simulation at $z$ = 9.3. The thick, black line is the analytic Sheth-Tormen mass function. The distribution of halo masses matches well with analytical expectations above 10$^{5}$ M$_{\odot}$, where halos are captured by 50 particles. \textbf{The vertical blue line indicates our halo mass resolution limit of 10$^{5}$ M$_{\odot}$.}}
    \label{fig:hmf}
\end{figure}

@@

%====================================================================
\subsection{Population III host halo masses}
%====================================================================
 To determine the host halos of new Pop III stars, we use the halo finding code, \rockstar{} \citep{rockstar} \textbf{(standing for Robust Overdensity Calculation using K-Space Topologically Adaptive Refinement), a friends-of-friends like algorithm which adaptively refines in six phase-space dimensions and one time dimension}. The mass determined by \rockstar{} is the dark matter mass, so to determine the total mass of the halos, we multiply the masses by $\Omega_{\rm M}/\Omega_{\rm DM}$. The total mass is subsequently used in our analysis. We also take the center and virial radius from the \rockstar{} halo catalog. A small percentage of Pop III stars do not form in a halo identified by \rockstar{}. Assuming that Pop III stars must form in collapsed halos, we calculate the virial mass and radius of a halo centered on the Pop III star directly from the simulation data by determining the sphere that contains an average overdensity of 200 $\rho_{\rm c}$. We calculate the LW intensity impinging on each halo by summing up the contribution coming from each radiating star particle outside the virial radius of the host halo. This is then added to the constant LW background implemented in the simulation (Eq. \ref{LWbg}). In this work, we only analyze halos below the atomic cooling limit.

### 2019-08-05 referee.tex
Fixed some formatting. Added notes to referee report and fixed a few different things within the paper (all in bold).

-Resolution testing

\begin{referee}
2) The authors select their haloes by the halo finding algorithm Rockstar. They should add a bit of information on that code (is it fof or subfind-like?). Also, one of their main points is to compare their work to Machacek+01. Machacek+01, however, use a different halo definition (spherical around the halo centre going out to a radius of 200 times the mean density of the Universe). How do these different halo masses compare?
Further, it is not entirely clear why atomic cooling haloes are excluded in this analysis and if they are excluded throughout the whole paper. The exclusion of atomic cooling haloes artificially shifts the average halo mass to lower values.
In addition, the authors should state the fraction of haloes that have Pop III stars forming outside the halo.
The baryon fraction varies with redshift. Please give an estimate by how much this influences the halo masses used here.
\end{referee}

-address concern about Rockstar details, and Machacek halo definition
-exclusion of atomic cooling haloes - we excluded these because of a bug in our code... how to explain this without simply saying that?
-fraction of halos that halo Pop III stars forming outside the haloes: I'm not sure I will be able to tell exactly how many are not in halos. I need to look back at the data to see if I will be able to pick out which halos I artificially drew a halo around. 
-shall I look at the baryon fraction as a function of time to see how much it varies?

\begin{referee}
Minor comments:
The authors should state at one point in the paper for which properties they use comoving or physical units.
\end{referee}
-We use physical units throughout the whole paper, apart from when we discuss the boxsize. I will make this more clear.

@@

\begin{referee}
The authors name a few processes for suppressing Pop III star formation. For completeness, they should also mention eg X-ray heating (citing eg Jeon+14). 
\end{referee}
-I will add this detail

\begin{referee}
The authors cite the well-known study from Machacek+01. However, a more recent result would be from the study by Visbal+14, giving a lower mass limit. Since the authors find a lower mass limit than Machacek+01, it would be interesting to see how this compares to the lower mass limit from the more modern paper.
\end{referee}
-I will check out Visbal+ 14 (I think I have looked at this paper, will double check)

\begin{referee}
Methods:
In Wise+12, three different environment regions are picked. Please provide more description on the rare peak simulation here and discuss why this one is appropriate in this context.
\end{referee}
-I will look at that simulation again, if John could provide some insight here that would be nice.

@@

\begin{referee}
Do the authors set fixed mass limits to the IMF and if yes, which are they? If not, which is the highest / lowest mass they find in their random sampling? And which fraction of the Pop III stars in the simulation end up as type II SNe / BHs / PISN?
\end{referee}
-We do set fixed mass limits, between 1 and 300 Msol. I will add this detail. 
-What is the best way to get this information? Shall I look at all the Pop III stars at the end of the simulation and look at their masses? I should be able to get a fraction of these endings from that, correct?

@@

\begin{referee}
In Figure 1, please mention in the caption or at the top / bottom of the left axis that a shielding factor of 1 corresponds to no shielding and a shielding factor of 0 corresponds to high shielding, as these nomenclatures are not immediately understandable for the reader.
\end{referee}
-This is a good point, that is confusing. I will add this.

\begin{referee}
The analytic calculation of the shielding factor is a nice property of the paper. However, assuming a constant H2 fraction seems not appropriate. Please repeat this exercise for a more realistic halo profile, which could be found in the simulations or in the literature, eg by Machacek+01 themselves.
\end{referee}
-I will check out a more realistic H2 fraction for a halo and try to get a nice plot out of that.

\begin{referee}
Results:
Figure 2 is quite empty, it would be useful to include the halo mass function of the earliest Pop III formation redshift with its analytic prediction. In addition, the authors could provide a line or an arrow to show their halo mass resolution limit.
\end{referee}
-Figure 2 is empty. I will add a vertical line at 10^5 Msol to show our resolution limit. As far as the HMF of the earliest Pop III formation redshift, is this referring to the HMF at the time when Pop III stars first begin to form, so at about z=27, what is the HMF at that time and how does it compare with Sheth-Tormen? I will plot this up.

\begin{referee}
Figure 3 looks very nice. It would be helpful if an additional line that only provides the global LWBG (in units of J21) was included.
\end{referee}
-Woohoo! Figure 3 does look nice! I will add the global LWBG, hopefully it looks alright. I am fairly certain we did this and for some reason it all didn't seem to add up. I will try this again.

@@

\begin{referee}
Figures 7 and 8 show very nice histograms as an additional panel on the right. I suggest to include a histogram on the spread of the creation time in Figure 9 similar to Figures 7 and 8.
\end{referee}
-I will add a histogram on the spread of the creation time in Figure 9.

@@

\begin{referee}
One of the most extensive studies on over 1000 minihaloes was performed by Hirano+15, who also include a Lyman-Werner background. Their work should be put in comparison as well.
\end{referee}
-I will check out this paper, and see what can be added.

\begin{referee}
In Figure 10, the result from Yoshida+03 is hard to see, maybe use a different symbol / colour.
\end{referee}
-I will adjust this symbol.

\begin{referee}
Since the authors study star formation in minihaloes, it would be appropriate to also mention Kitayama+04 and Schauer+15 for the Lyman-Werner escape fractions from minihaloes.
\end{referee}
-I will check out these papers and add accordingly.

\begin{referee}
The authors mention a few times that streaming velocities are another pathway of suppressing Pop III formation (also in the introduction). While they cite some older work, they however fail to compare to more recent work eg by Hirano+18 or Schauer+19 (who do perform a systematic study not only based on a few, single haloes).
\end{referee}
-I will check out these papers as well, and see if I can compare.

>>>>

-Resolution testing
-Ran original sim to see if it worked, it did run, but only reached until z=22. The simulation stopped because of a stopping parameter, StopFirstTimeAtLevel.

\begin{referee}
2) The authors select their haloes by the halo finding algorithm Rockstar. They 
should add a bit of information on that code (is it fof or subfind-like?). Also,
 one of their main points is to compare their work to Machacek+01. Machacek+01, 
 however, use a different halo definition (spherical around the halo centre 
 going out to a radius of 200 times the mean density of the Universe). How do 
 these different halo masses compare?

Further, it is not entirely clear why atomic cooling haloes are excluded in this
 analysis and if they are excluded throughout the whole paper. The exclusion of 
 atomic cooling haloes artificially shifts the average halo mass to lower 
 values.

In addition, the authors should state the fraction of haloes that have Pop III 
stars forming outside the halo. The baryon fraction varies with redshift. Please
 give an estimate by how much this influences the halo masses used here.
\end{referee}

-Added a bit about what Rockstar is. 
-Machacek halo definition - how should I quantify this difference?

-Exclusion of atomic cooling haloes - we excluded these because of a bug in our code... how to explain this without simply saying that?

-Fraction of halos that halo Pop III stars forming outside the halos: I'm not sure I will be able to tell exactly how many are not in halos. I need to look back at the data to see if I will be able to pick out which halos I artificially drew a halo around. 
-Shall I look at the baryon fraction as a function of time to see how much it varies?

\begin{referee}
Minor comments:
The authors should state at one point in the paper for which properties they use
 comoving or physical units.
\end{referee}
-Clarified this at the end of the second paragraph in the Simulation Setup section.

@@

\begin{referee}
The authors name a few processes for suppressing Pop III star formation. For 
completeness, they should also mention eg X-ray heating (citing eg Jeon+14). 
\end{referee}
-I will add this detail (https://arxiv.org/pdf/1310.7944.pdf)
-Jeon+14 seems to show that X-ray heating does not affect star formation, but 
the clumping factor is reduced, and thus makes it easier to keep the reionized 
gas ionized.

\begin{referee}
The authors cite the well-known study from Machacek+01. However, a more recent 
result would be from the study by Visbal+14, giving a lower mass limit. Since 
the authors find a lower mass limit than Machacek+01, it would be interesting to
 see how this compares to the lower mass limit from the more modern paper.
\end{referee}
-Visbal+14 provides an equation for Mcrit (eqn. 4), I have plotted this previously. This lines up fairly well with our data, with a bit of scatter.  (https://arxiv.org/pdf/1402.0882.pdf)

\begin{referee}
Methods:
In Wise+12, three different environment regions are picked. Please provide more 
description on the rare peak simulation here and discuss why this one is 
appropriate in this context.
\end{referee}
-RP stands for radiation pressure in Wise+12 (correct?), not rare peak. The simulation is the same size as ours, and contains similar physics, including the radiation pressure and soft UV background, which is why we reference it here. The differences are the inclusion of H2 self-shielding and updated cosmo parameters.

@@

\begin{referee}
Do the authors set fixed mass limits to the IMF and if yes, which are they? 
If not, which is the highest / lowest mass they find in their random sampling? 
And which fraction of the Pop III stars in the simulation end up as type II 
SNe / BHs / PISN?
\end{referee}
-We do set fixed mass limits, between 1 and 300 Msol. I added this detail. 
-What is the best way to get this information? Shall I look at all the Pop III stars at the end of the simulation and look at their masses? I should be able to get a fraction of these endings from that, correct?

@@

\begin{referee}
In Figure 1, please mention in the caption or at the top / bottom of the left 
axis that a shielding factor of 1 corresponds to no shielding and a shielding 
factor of 0 corresponds to high shielding, as these nomenclatures are not 
immediately understandable for the reader.
\end{referee}
-This detail has been added.

\begin{referee}
The analytic calculation of the shielding factor is a nice property of the 
paper. However, assuming a constant H2 fraction seems not appropriate. Please 
repeat this exercise for a more realistic halo profile, which could be found in 
the simulations or in the literature, eg by Machacek+01 themselves.
\end{referee}
-I will check out a more realistic H2 fraction for a halo and try to get a nice plot out of that.
-Any suggestions from John for a more realistic halo profile?

\begin{referee}
Results:
Figure 2 is quite empty, it would be useful to include the halo mass function of
 the earliest Pop III formation redshift with its analytic prediction. In 
 addition, the authors could provide a line or an arrow to show their halo mass 
 resolution limit.
\end{referee}
-As far as the HMF of the earliest Pop III formation redshift, is this referring to the HMF at the time when Pop III stars first begin to form, so at about z=27, what is the HMF at that time and how does it compare with Sheth-Tormen? I will plot this up.

-Added a vertical line at $10^5 M_{\odot}$. Detail also added to caption.
-Tried plotting the HMF of the earliest Pop III formation redshift, but I need to collect the HMF of that dataset -- having troubles remembering how this was produced

\begin{referee}
Figure 3 looks very nice. It would be helpful if an additional line that only 
provides the global LWBG (in units of J21) was included.
\end{referee}
-Woohoo! Figure 3 does look nice!
-I tried adding the global LWBG, but this makes our results seem funny, because we are getting the average J21 weighted by cell mass lying below the global LWBG. Thoughts about this?

@@

\begin{referee}
Figures 7 and 8 show very nice histograms as an additional panel on the right. 
I suggest to include a histogram on the spread of the creation time in Figure 9 
similar to Figures 7 and 8.
\end{referee}
-Added the histogram to Figure 9.

@@

\begin{referee}
One of the most extensive studies on over 1000 minihaloes was performed by 
Hirano+15, who also include a Lyman-Werner background. Their work should be put 
in comparison as well.
\end{referee}
Hirano+15 is looking at the distribution of stellar masses to calculate the IMF of p3 stars, which is why we quote them in the introduction. 
-Their Fig. 3 shows the number of halos versus virial mass that host primordial star-forming clouds. 
-Using their eqn. 6 they find Mvir = $2.1 \times 10^5$ at z=30, and Mvir = $9.9 \times 10^5$ at z=10, which they say is in good agreement with their Fig. 3. 
(https://arxiv.org/pdf/1501.01630.pdf)

\begin{referee}
In Figure 10, the result from Yoshida+03 is hard to see, maybe use a different 
symbol / colour.
\end{referee}
-Adjusted. Changed the symbol to a pentagon and the outline color black. Okay?

\begin{referee}
Since the authors study star formation in minihaloes, it would be appropriate to
 also mention Kitayama+04 and Schauer+15 for the Lyman-Werner escape fractions 
 from minihaloes.
\end{referee}
-Kitayama+04: HII regions are governed by an initial slow and weak D-type ionization front followed by a rapid R-type front. For there to be complete ionization, the transition between the types is critical. They find that in small mass halos, less than $10^6 M_{\odot}$, this critical transition takes place fairly rapidly, allowing for high escape fractions of greater than 80\%. In larger mass halos, the ionization front remains as a D-type, and the escape fraction is essentially zero. They use a simulation to study these HII regions in similar mass halos as ours and within a similar redshift range. Their Fig. 4 shows the ionizing and LW escape fractions as a function of halo mass for different stellar masses. (https://arxiv.org/pdf/astro-ph/0406280.pdf)

-Schauer+15: They run simulations of LW escape fractions from P3 stars in $10^5 - 10^7 M_{\odot}$ halos. They find escape fractions from 0-85\%, and that H2 shielding by neutral hydrogen also decreases escape fractions compared to escape fractions predicted by only H2 self-shielding. Fig.3 shows escape fractions for different models and different P3 stellar masses. When H2 shielding by neutral hydrogen is not included, escape fractions are overpredicted. (https://arxiv.org/pdf/1506.04796.pdf)

\begin{referee}
The authors mention a few times that streaming velocities are another pathway of
 suppressing Pop III formation (also in the introduction). While they cite some 
 older work, they however fail to compare to more recent work eg by Hirano+18 or
  Schauer+19 (who do perform a systematic study not only based on a few, single 
  haloes).
\end{referee}
-Hirano+18: Investigate the collapse of primordial gas to form P3 stars when the energy transfer between baryon and dark matter fluids (due to non-gravitational scattering) is included.
